---
title: 'Nuclear Power Attitudes: Network Analyses'
author: "Kraft, B., Chambon, M., & Kraft, P."
date: "2025-05-22"
output: html_document
---

```{r Setup & libraries, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(foreign)
library(dplyr)
library(qgraph)
library(bootnet)
library(igraph)
library(kableExtra)
library(NetworkComparisonTest)
data_path <- here("data")
output_path  <- here("output")
```

Data: <a href="https://osf.io/s7t2n/" title="osf.io/s7t2n">https://osf.io/s7t2n/</a>

R version:
```{r PackageVersions}
sessionInfo()
```

```{r load data, include=FALSE}
data <- read.spss(
  here(data_path, "OSF_data_attitudes_towards_nuclear_power_12082024.sav"),
  to.data.frame = TRUE,
  na.strings = "999",
  stringsAsFactors = FALSE
)
```
# 1. General population network
```{r Generate Data Frames for Network Analyses, include=FALSE}
# General population network
data_network1 <- data %>% select(accident, depend, solvewa, expensiv, stable, terrwar, gases,
                                 locburd, price, birds, waste, employ, water)
```
```{r Estimate general population network and communities, include=FALSE}
network_general_public <- estimateNetwork(data_network1, default = "EBICglasso", corMethod = "cor_auto", missing = "pairwise")
network_general_public_centrality_plot <- centralityPlot(network_general_public, include = c("Strength"), scale = "raw")
network_general_public_igraph <- graph_from_adjacency_matrix(abs(network_general_public$graph), 'undirected', weighted = TRUE, add.colnames = FALSE)
network_general_public_communities <- cluster_walktrap(network_general_public_igraph)
network_general_public_plot <- plot(network_general_public, theme = "colorblind", groups = communities(network_general_public_communities))
network_general_public_plot_layout <- network_general_public_plot$layout
```

```{r Naming nodes and communities, include=FALSE}
# Naming nodes and communities
network_general_public_communities_nodenames <- list("Nuclear power may cause accidents", "Less dependent on other countries",
                                                       "Solves problems w/ global warming", "Are expensive to build", "Provides a stable source of energy",
                                                       "Exposed for terror and war", "Prod. power w/o greenhouse gases",
                                                       "Represents a burden to local comm.", "Keeps price on electricity low", 
                                                       "Takes little space & is good for animals", "Produces problematic waste", "Contributes to employment in Norway",
                                                       "Produces hot water that can be used")
network_general_public_communities_grouplabels <- list("Materialist" = c(2,9,3,12), "Environment" = c(7,5,1,10,13), "Burden" = c(8,6,11,4))
```
## 1.2 Network
```{r Plotting network, echo=FALSE}
plot(network_general_public, theme = "colorblind", groups = network_general_public_communities_grouplabels, 
     nodeNames = network_general_public_communities_nodenames ,title = "General population network", legend.cex = 0.3,  width = 14,  height = 10)
```

## 1.3 Centrality indices
```{r Plotting centrality indices, echo=FALSE}
print(network_general_public_centrality_plot)
```


```{r Write SVGs, include=FALSE}
svg(filename = here("output", "general_population", "general_population_network.svg"),  width = 14,  height = 10)
plot(network_general_public, theme = "colorblind", groups = network_general_public_communities_grouplabels, 
     nodeNames = network_general_public_communities_nodenames ,title = "General population network", legend.cex = 0.5)
dev.off()
svg(filename = here("output", "general_population", "centrality.svg"), width = 6, height = 12)
print(network_general_public_centrality_plot)
dev.off()
```

## 1.3 Statistics
**Table S1.**  
Edge weights and centrality indices for the general population network
```{r Extract edge weights, echo=FALSE}
# Extract edge weights and centrality indices
edges_network_general_public <- network_general_public$graph
edges_network_general_public <- as.data.frame(edges_network_general_public)
edges_network_general_public_half <- edges_network_general_public
edges_network_general_public_half[lower.tri(as.matrix(edges_network_general_public_half))] <- NA
edges_network_general_public_half[edges_network_general_public_half == 0] <- NA
edges_network_general_public_half <- round(edges_network_general_public_half, 2)
# Write table
edges_network_general_public_half <- edges_network_general_public_half[-nrow(edges_network_general_public_half), ]
edges_network_general_public_half[] <- lapply(edges_network_general_public_half, function(x) {
  ifelse(is.na(x), "", x)
})
edges_network_general_public_half %>%
  kbl(align = "c") %>%
  kable_classic(full_width = TRUE)
```
**Table S2**  
Centrality indices (strength) for the general population network
```{r Unstandardized centrality indices, echo=FALSE}
  
## Unstandardized centrality measures
centTable_network_general_public<-centralityTable(network_general_public, standardized = FALSE)
centTable_strength <- centTable_network_general_public[centTable_network_general_public$measure == "Strength", ]
centTable_strength$value <- round(centTable_strength$value, 1)
centTable_strength <- centTable_strength[, -c(1, 2, 4)]
rownames(centTable_strength) <- NULL
colnames(centTable_strength)[1] <- "Node"
colnames(centTable_strength)[2] <- "Strength"
centTable_strength %>%
  kbl(
    align = "c",
  ) %>%
  kable_classic(full_width = TRUE) %>%
  add_header_above(c(" " = ncol(centTable_strength)))  # prevents kableExtra from dropping APA style
```
## 1.4 Bootstrap analyses
### 1.4.1 Edge accuracy
```{r Statistical accuracy & signifcance, include=FALSE, cache=TRUE}
set.seed(123)
Bootnet_network_general_public <- bootnet(data_network1, nBoots = 1000, corMethod = "cor_auto", 
                              default = "EBICglasso", missing = "pairwise")

```
```{r Plot edge accuracy for the general population network, echo=FALSE, fig.height=10}

plot(Bootnet_network_general_public, labels = TRUE, order = "sample", res=150)

```


### 1.4.2 Edge differences tests
```{r Edge differences, echo=FALSE}
plot(Bootnet_network_general_public, statistics = "edge", plot = "difference", onlyNonZero = TRUE)
Bootnet_network_general_public[["sampleTable"]][["id"]]
```
```{r Difference test, echo=FALSE, eval=FALSE}
differenceTest(Bootnet_network_general_public, "depend--stable" ,"stable--gases", "edge")
```

### 1.4.3 Centrality stability and significant differences
```{r Centrality stability estimation, include=FALSE, cache=TRUE}
set.seed(123)
CentStability_network_general_public <- bootnet(data_network1, nBoots = 1000, corMethod = "cor_auto",
                                    default = "EBICglasso", type = "case",
                                    statistics = c("Strength"), missing = "pairwise")
```
```{r Plot centrality stability, echo=FALSE}
corStability(CentStability_network_general_public)
plot(CentStability_network_general_public, subsetRange = c(100,50), statistics = c("Strength"))
plot(CentStability_network_general_public, "Strength", perNode = TRUE, subsetRange = c(100,50))
plot(Bootnet_network_general_public, statistics = c("Strength"), plot = "difference")
```


# 2. Comparing positive and negative attitude networks
```{r Loading data per group, echo=FALSE}
#Positive attitudes network
data_network2_pos_att <- data %>%  filter(ATT3GROU == 3) %>%
  select(accident, depend, solvewa, expensiv, stable, terrwar, gases, 
         locburd, price, birds, waste, employ, water)
#Negative attitudes network
data_network3_neg_att <- data %>%  filter(ATT3GROU == 1 ) %>%
  select(accident, depend, solvewa, expensiv, stable, terrwar, gases, 
         locburd, price, birds, waste, employ, water)
```
## 2.1 Positive attitudes network
### 2.1.1 Network
```{r Positive attitudes network & centrality stability, include=FALSE}
network2_pos_att <- estimateNetwork(data_network2_pos_att, default = "EBICglasso", corMethod = "cor_auto", missing = "pairwise")
```
```{r Plot positive attitudes network, echo=FALSE}
plot(network2_pos_att, theme = "colorblind", groups = network_general_public_communities_grouplabels, 
     nodeNames = network_general_public_communities_nodenames ,title = "Positive population network", legend.cex = 0.3,  width = 14,  height = 10)
```

### 2.1.2 Centrality stability
```{r Estimate centrality stability for the positive attitudes network, include=FALSE, cache=TRUE}
set.seed(123)
CentStability_network2_pos_att <- bootnet(data_network2_pos_att, nBoots = 1000, corMethod = "cor_auto",
                                                default = "EBICglasso", type = "case",
                                                statistics = c("Strength"), missing = "pairwise")
```
```{r Print CS-coefficient for the positive attitudes network, echo=FALSE}
corStability(CentStability_network2_pos_att)
```
### 2.1.3 Statistics
**Table S3.**  
Edge weights and centrality indices for the positive attitudes network
```{r Extract edge weights for the positive attitudes network, echo=FALSE}
# Extract edge weights and centrality indices
edges_network_positive_attitudes <- network2_pos_att$graph
edges_network_positive_attitudes <- as.data.frame(edges_network_positive_attitudes)
edges_network_positive_attitudes_half <- edges_network_positive_attitudes
edges_network_positive_attitudes_half[lower.tri(as.matrix(edges_network_positive_attitudes))] <- NA
edges_network_positive_attitudes_half[edges_network_positive_attitudes_half == 0] <- NA
edges_network_positive_attitudes_half <- round(edges_network_positive_attitudes_half, 2)
# Write table
edges_network_positive_attitudes_half <- edges_network_positive_attitudes_half[-nrow(edges_network_positive_attitudes_half), ]
edges_network_positive_attitudes_half[] <- lapply(edges_network_positive_attitudes_half, function(x) {
  ifelse(is.na(x), "", x)
})
edges_network_positive_attitudes_half %>%
  kbl(align = "c") %>%
  kable_classic(full_width = TRUE)
```
**Table S4**  
Centrality indices (strength) for the positive attitudes network
```{r Unstandardized centrality indices for the positive attitudes network, echo=FALSE}
  ## Unstandardized centrality measures
centTable_network_positive_attitudes<-centralityTable(network2_pos_att, standardized = FALSE)
centTable_network_positive_attitudes_strength <- centTable_network_positive_attitudes[centTable_network_positive_attitudes$measure == "Strength", ]
centTable_network_positive_attitudes_strength$value <- round(centTable_network_positive_attitudes_strength$value, 1)
centTable_network_positive_attitudes_strength <- centTable_network_positive_attitudes_strength[, -c(1, 2, 4)]
rownames(centTable_network_positive_attitudes_strength) <- NULL
colnames(centTable_network_positive_attitudes_strength)[1] <- "Node"
colnames(centTable_network_positive_attitudes_strength)[2] <- "Strength"
centTable_network_positive_attitudes_strength %>%
  kbl(
    align = "c",
  ) %>%
  kable_classic(full_width = TRUE) %>%
  add_header_above(c(" " = ncol(centTable_network_positive_attitudes_strength)))  # prevents kableExtra from dropping APA style
```

### 2.1.4 Bootstrap analyses
#### 2.1.4.1 Edge accuracy
```{r Statistical accuracy & signifcance for the positive attitudes network, include=FALSE, cache=TRUE}
set.seed(123)
Bootnet_positive_attitudes_network <- bootnet(data_network2_pos_att, nBoots = 1000, corMethod = "cor_auto", 
                              default = "EBICglasso", missing = "pairwise")

```
```{r Plot edge accuracy for  the positive attitudes network, echo=FALSE, fig.height=10}

plot(Bootnet_positive_attitudes_network, labels = TRUE, order = "sample", res=150)

```


#### 2.1.4.2 Edge differences tests
```{r Edge differences for the positive attitudes network, echo=FALSE}
plot(Bootnet_positive_attitudes_network, statistics = "edge", plot = "difference", onlyNonZero = TRUE)
Bootnet_positive_attitudes_network[["sampleTable"]][["id"]]
```
```{r Difference test for the positive attitudes network, echo=FALSE, eval=FALSE}
differenceTest(Bootnet_positive_attitudes_network, "depend--stable" ,"stable--gases", "edge")
```

#### 2.1.4.3 Centrality: significant differences
```{r Centrality: significant differences for the positive attitudes network, include=FALSE, cache=TRUE}
set.seed(123)
CentStability_positive_attitudes_network <- bootnet(data_network2_pos_att, nBoots = 1000, corMethod = "cor_auto",
                                    default = "EBICglasso", type = "case",
                                    statistics = c("Strength"), missing = "pairwise")
```
```{r Plot centrality stabilityfor the positive attitudes network, echo=FALSE}
corStability(CentStability_positive_attitudes_network)
plot(CentStability_positive_attitudes_network, subsetRange = c(100,50), statistics = c("Strength"))
plot(CentStability_positive_attitudes_network, "Strength", perNode = TRUE, subsetRange = c(100,50))
plot(Bootnet_positive_attitudes_network, statistics = c("Strength"), plot = "difference")
```





## 2.2 Negative attitudes network
### 2.2.1 Network
```{r Negative attitudes network & centrality stability, include=FALSE}
network3_neg_att <- estimateNetwork(data_network3_neg_att, default = "EBICglasso", corMethod = "cor_auto", missing = "pairwise")
```
```{r Plot negative attitudes network, echo=FALSE}
plot(network3_neg_att, theme = "colorblind", groups = network_general_public_communities_grouplabels, 
     nodeNames = network_general_public_communities_nodenames ,title = "Negative population network", legend.cex = 0.3,  width = 14,  height = 10)
```

### 2.2.2 Centrality stability
```{r Estimate centrality stability for the negative attitudes network, include=FALSE, cache=TRUE}
set.seed(123)
CentStability_network3_neg_att <- bootnet(network3_neg_att, nBoots = 1000, corMethod = "cor_auto",
                                                default = "EBICglasso", type = "case",
                                                statistics = c("Strength"), missing = "pairwise")
```
```{r Print CS-coefficient for the negative attitudes network, echo=FALSE}
corStability(CentStability_network3_neg_att)
```

## 2.3 Undecided/negative attitudes network
### 2.3.1 Network
```{r Loading undecided/negative data, echo=FALSE}
data_network4_undec_neg_att <- data %>%  filter(ATT3GROU < 3 ) %>%
  select(accident, depend, solvewa, expensiv, stable, terrwar, gases, 
         locburd, price, birds, waste, employ, water)
```
```{r Undecided/negative attitudes network & centrality stability, include=FALSE}
network4_undec_neg_att <- estimateNetwork(data_network4_undec_neg_att, default = "EBICglasso", corMethod = "cor_auto", missing = "pairwise")
```
```{r Plot undecided/negative attitudes network, echo=FALSE}
plot(network4_undec_neg_att, theme = "colorblind", groups = network_general_public_communities_grouplabels, 
     nodeNames = network_general_public_communities_nodenames ,title = "Undecided/negative network", legend.cex = 0.3,  width = 14,  height = 10)
```

### 2.3.2 Centrality stability
```{r Estimate centrality stability for the undecided/negative attitudes network, include=FALSE, cache=TRUE}
set.seed(123)
CentStability_network4_undec_neg_att <- bootnet(data_network4_undec_neg_att, nBoots = 1000, corMethod = "cor_auto",
                                                default = "EBICglasso", type = "case",
                                                statistics = c("Strength"), missing = "pairwise")
```
```{r Print CS-coefficient for the undecided/negative attitudes network, echo=FALSE}
corStability(CentStability_network4_undec_neg_att)
```


### 2.3.3 Statistics
**Table S6.**  
Edge weights and centrality indices for the undecided/negative attitudes network
```{r Extract edge weights for the undecided/negative attitudes network, echo=FALSE}
# Extract edge weights and centrality indices
edges_network_undec_neg_attitudes <- network4_undec_neg_att$graph
edges_network_undec_neg_attitudes <- as.data.frame(edges_network_undec_neg_attitudes)
edges_network_undec_neg_attitudes_half <- edges_network_undec_neg_attitudes
edges_network_undec_neg_attitudes_half[lower.tri(as.matrix(edges_network_undec_neg_attitudes_half))] <- NA
edges_network_undec_neg_attitudes_half[edges_network_undec_neg_attitudes_half == 0] <- NA
edges_network_undec_neg_attitudes_half <- round(edges_network_undec_neg_attitudes_half, 2)
# Write table
edges_network_undec_neg_attitudes_half <- edges_network_undec_neg_attitudes_half[-nrow(edges_network_undec_neg_attitudes_half), ]
edges_network_undec_neg_attitudes_half[] <- lapply(edges_network_undec_neg_attitudes_half, function(x) {
  ifelse(is.na(x), "", x)
})
edges_network_undec_neg_attitudes_half %>%
  kbl(align = "c") %>%
  kable_classic(full_width = TRUE)
```
**Table S7**  
Centrality indices (strength) for the undecided/negative attitudes network
```{r Unstandardized centrality indices for the undecided/negative attitudes network, echo=FALSE}
  ## Unstandardized centrality measures
centTable_network_undec_neg_attitudes<-centralityTable(network4_undec_neg_att, standardized = FALSE)
centTable_network_undec_neg_attitudes_strength <- centTable_network_undec_neg_attitudes[centTable_network_undec_neg_attitudes$measure == "Strength", ]
centTable_network_undec_neg_attitudes_strength$value <- round(centTable_network_undec_neg_attitudes_strength$value, 1)
centTable_network_undec_neg_attitudes_strength <- centTable_network_undec_neg_attitudes_strength[, -c(1, 2, 4)]
rownames(centTable_network_undec_neg_attitudes_strength) <- NULL
colnames(centTable_network_undec_neg_attitudes_strength)[1] <- "Node"
colnames(centTable_network_undec_neg_attitudes_strength)[2] <- "Strength"
centTable_network_undec_neg_attitudes_strength %>%
  kbl(
    align = "c",
  ) %>%
  kable_classic(full_width = TRUE) %>%
  add_header_above(c(" " = ncol(centTable_network_undec_neg_attitudes_strength)))  # prevents kableExtra from dropping APA style
```

### 2.4 Bootstrap analyses
#### 2.4.1 Edge accuracy
```{r Statistical accuracy & signifcance for the undecided/negative attitudes network, include=FALSE, cache=TRUE}
set.seed(123)
Bootnet_undec_neg_attitudes_network <- bootnet(data_network4_undec_neg_att, nBoots = 1000, corMethod = "cor_auto", 
                              default = "EBICglasso", missing = "pairwise")

```
```{r Plot edge accuracy for  the undecided/negative attitudes network, echo=FALSE, fig.height=10}

plot(Bootnet_undec_neg_attitudes_network, labels = TRUE, order = "sample", res=150)

```


#### 2.4.2 Edge differences tests
```{r Edge differences for the undecided/negative attitudes network, echo=FALSE}
plot(Bootnet_undec_neg_attitudes_network, statistics = "edge", plot = "difference", onlyNonZero = TRUE)
Bootnet_undec_neg_attitudes_network[["sampleTable"]][["id"]]
```
```{r Difference test for the undecided/negative attitudes network, echo=FALSE, eval=FALSE}
differenceTest(Bootnet_undec_neg_attitudes_network, "depend--stable" ,"stable--gases", "edge")
```

#### 2.4.3 Centrality: significant differences
```{r Centrality significant differences for the undecided/negative attitudes network, include=FALSE, cache=TRUE}
set.seed(123)
CentStability_undec_neg_attitudes_network <- bootnet(data_network4_undec_neg_att, nBoots = 1000, corMethod = "cor_auto",
                                    default = "EBICglasso", type = "case",
                                    statistics = c("Strength"), missing = "pairwise")
```
```{r Plot centrality stabilityfor the undecided/negative attitudes network, echo=FALSE}
corStability(CentStability_undec_neg_attitudes_network)
plot(CentStability_undec_neg_attitudes_network, subsetRange = c(100,50), statistics = c("Strength"))
plot(CentStability_undec_neg_attitudes_network, "Strength", perNode = TRUE, subsetRange = c(100,50))
plot(Bootnet_undec_neg_attitudes_network, statistics = c("Strength"), plot = "difference")
```


## 2.5 Statistical comparison of the positive and undecided/negative attitude networks
```{r Statistical comparison of positive vs. undecided/negative network, include=FALSE, cache=TRUE}
set.seed(123)
comparing_pos_neg_att_nct_results <- NCT(
  network2_pos_att, network4_undec_neg_att, it = 1000, test.edges = TRUE, test.centrality = TRUE, centrality = c("strength"))
```
```{r Results from statistical comparison of positive vs. undecided/negative network, echo=FALSE}
comparing_pos_neg_att_nct_results
comparing_pos_neg_att_nct_results$einv.pvals[which(comparing_pos_neg_att_nct_results$einv.pvals[,3] < 0.05), ]
```


